---
title: "Shifting car travel to active modes to improve population health and achieve transport goals: a simulation study"
author: Belen Zapata-Diomedi, Alan Both, Ali Abbas, James Woodcock, Annette Kroen, Melanie Davern, Lucy Gunn
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: column
    source_code: embed
    storyboard: true
    navbar:
runtime: shiny
---

```{r global, include=FALSE}

library(dplyr)
library(knitr)
library(shinyWidgets)
library(tidyverse)
library(flexdashboard)
library(shiny)


# Load data ---------------------------------------------------------------
output_transport_modes <- read_csv('output_transport_modes.csv')

diseaseLevels <- c("ishd","strk",
                   "brsc","carc","tbalc","utrc","mltm", "chml","stmc", "lvrc","hanc",
                   "dmt2", 
                   "dprd", 
                   "adaod")
diseaseLabels <- c("Ischemic heart disease","Stroke",
                   "Breast cancer","Colon cancer","Lung cancer","Uterine cancer","Multiple myeloma", 
                   "Chronic myeloid leukemia", "Stomach cancer", "Liver cancer","Head and neck cancer",
                   "Diabetes type 2",
                   "Depression", 
                   "Alzheimers disease and other dementias")
```

```{r data}

### Create user inputs

scenario_choice <- unique(output_transport_modes$title1)
age_choice <- unique(output_transport_modes$age)
sex_choice <- unique(output_transport_modes$sex) 
purpose_choice <- unique(output_transport_modes$purpose)
```

# Scenarios {.sidebar}

Use the tabs to select which age and sex groups, distances travelled by car to be replaced by walking, cycling or a combination of both and type of trips.

```{r}
# Graph showing changes in transport mode share for selected scenarios

### selection Input based on columns age, sex and scenario (title1)
selectInput("scenario", label= "Scenario:",
            choices = scenario_choice,
            selected = scenario_choice[[1]])

selectInput("age", label= "Age group:",choices = age_choice,
            selected = age_choice[[1]])

selectInput("sex", label= "Sex:",
            choices = sex_choice,
            selected = sex_choice[[1]])

selectInput("purpose", label = "Purpose:",
            choices = purpose_choice,
            selected = purpose_choice[[1]])


```

# Transport mode share

```{r}

renderPlot({output_transport_modes %>%
                filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
   ggplot(aes(x=prop, y=mode, fill=scenario)) +
    geom_bar(stat="identity", position="dodge") +
    scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
    labs(x="Proportion of all trips") +
    geom_text(aes(label=paste0(round(prop*100,1),"%")),
              position=position_dodge(width=0.9),
              hjust=-0.05, vjust=0.5, size=4) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 5L),limits=c(0,.8)) +
    theme_classic() +
   theme(plot.title=element_text(size=16, hjust = 0.5),
          axis.text=element_text(size=14),
          axis.title=element_text(size=14),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=14),
          legend.key = element_blank(),
          strip.background = element_blank(),
          strip.text = element_text(size=14))})


```

# Incidence

### Accumulated number of diseases prevented over the life course of the modelled population.

```{r}

renderPlot({as_tibble(output_diseases_change) %>%
    mutate(purpose = case_when(grepl("all", scen) ~ "all",
                               grepl("commuting", scen) ~ "commuting")) %>%
     filter(measure=="inc.num") %>%
    # somewhere in the scripts tablc has been mislabeled tbalct
    mutate(disease=ifelse(disease=="tbalct","tbalc",disease)) %>%
    mutate(disease=factor(disease, levels=diseaseLevels, labels=diseaseLabels)) %>%
    mutate(across(median:percentile975, round, digits=3)) %>%
    dplyr::select(disease,median,percentile025,percentile975, scen, title1, age, sex, purpose, title1) %>% mutate(age=ifelse(age=="all", "all age groups", age),
           sex=ifelse(sex=="all", "females and males", sex), 
           sex=ifelse(sex=="male", "males", sex), 
           sex=ifelse(sex=="female", "females", sex)) %>%
                filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
                  ggplot(aes(x=median, y=disease)) +
    geom_bar(stat="identity", color=NA, fill="#E69F00", 
             position=position_dodge()) +
    labs(x="Change incidence") +
    geom_errorbar(aes(xmin=percentile025, xmax=percentile975), width=.2) +
    scale_x_continuous(labels = scales::comma, 
                       breaks = waiver(),
                       minor_breaks = NULL,
                       n.breaks = 10) +
    theme_classic() +
    theme(legend.position = "none") +
    theme(plot.title=element_text(size=16, hjust = 0.5),
          plot.subtitle = element_text(size=14, hjust = 0.5),
          axis.text=element_text(size=12),
          axis.title=element_text(size=12),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=12),
          axis.text.x = element_text(angle=90),
          legend.key = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(size=12))})
```

### Number of diseases prevented per year of simulation for the modelled population.

```{r}
renderPlot({as_tibble(output_df_agg_all) %>%
    mutate(purpose = case_when(grepl("all", scen) ~ "all",
                               grepl("commuting", scen) ~ "commuting")) %>%
     filter(measure=="inc.num") %>%
    # somewhere in the scripts tablc has been mislabeled tbalct
    mutate(disease=ifelse(disease=="tbalct","tbalc",disease)) %>%
    mutate(disease=factor(disease, levels=diseaseLevels, labels=diseaseLabels)) %>%
    mutate(across(median:percentile975, round, digits=3)) %>%
    dplyr::select(year, disease,median,percentile025,percentile975, scen, title1, age, sex, purpose, title1) %>% mutate(age=ifelse(age=="all", "all age groups", age),
           sex=ifelse(sex=="all", "females and males", sex), 
           sex=ifelse(sex=="male", "males", sex), 
           sex=ifelse(sex=="female", "females", sex)) %>%
                filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
  ggplot(aes(x=year, y=median, colour = disease)) +
    # geom_ribbon(aes(ymin=percentile025,ymax=percentile975),fill="#24C9AC",alpha=0.5) + # AUO teal, removed UI to graph all diseases in one graph
    geom_smooth(se=FALSE) + #or geom_line
    # facet_grid(rows=vars(disease), scales = "free_y") +
    scale_y_continuous(
      name = waiver(),
      breaks = waiver(),
      minor_breaks = NULL,
      n.breaks = 5,
      labels = waiver()) +
    scale_x_continuous(limits=c(0,85),breaks=seq(0,80,10), expand=c(0,0)) +
    labs(x="Years since scenario commenced", y="Incidence") +
    theme_classic() +
    theme(legend.position = "none") +
    theme(plot.title=element_text(size=16, hjust = 0.5),
          plot.subtitle = element_text(size=14, hjust = 0.5),
          axis.text=element_text(size=12),
          axis.title=element_text(size=12),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=12),
          legend.key = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(size=12)) + 
          geom_hline(yintercept = 0, color = "black")})

```

# Mortality

### Accumulated number of deaths prevented over the life course of the modelled population.

```{r}

renderPlot({as_tibble(output_diseases_change) %>%
    mutate(purpose = case_when(grepl("all", scen) ~ "all",
                               grepl("commuting", scen) ~ "commuting")) %>%
     filter(measure=="mx.num") %>%
    # somewhere in the scripts tablc has been mislabeled tbalct
    mutate(disease=ifelse(disease=="tbalct","tbalc",disease)) %>%
    mutate(disease=factor(disease, levels=diseaseLevels, labels=diseaseLabels)) %>%
    mutate(across(median:percentile975, round, digits=3)) %>%
    dplyr::select(disease,median,percentile025,percentile975, scen, title1, age, sex, purpose, title1) %>% mutate(age=ifelse(age=="all", "all age groups", age),
           sex=ifelse(sex=="all", "females and males", sex), 
           sex=ifelse(sex=="male", "males", sex), 
           sex=ifelse(sex=="female", "females", sex)) %>%
                filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
                  ggplot(aes(x=median, y=disease)) +
    geom_bar(stat="identity", color=NA, fill="#E69F00", 
             position=position_dodge()) +
    labs(x="Change mortality") +
    geom_errorbar(aes(xmin=percentile025, xmax=percentile975), width=.2) +
    scale_x_continuous(labels = scales::comma, 
                       breaks = waiver(),
                       minor_breaks = NULL,
                       n.breaks = 10) +
    theme_classic() +
    theme(legend.position = "none") +
    theme(plot.title=element_text(size=16, hjust = 0.5),
          plot.subtitle = element_text(size=14, hjust = 0.5),
          axis.text=element_text(size=12),
          axis.title=element_text(size=12),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=12),
          axis.text.x = element_text(angle=90),
          legend.key = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(size=12))})
    
```

### Number of deaths prevented per year of simulation for the modelled population.

```{r}
renderPlot({as_tibble(output_df_agg_all) %>%
    mutate(purpose = case_when(grepl("all", scen) ~ "all",
                               grepl("commuting", scen) ~ "commuting")) %>%
     filter(measure=="mx.num") %>%
    # somewhere in the scripts tablc has been mislabeled tbalct
    mutate(disease=ifelse(disease=="tbalct","tbalc",disease)) %>%
    mutate(disease=factor(disease, levels=diseaseLevels, labels=diseaseLabels)) %>%
    mutate(across(median:percentile975, round, digits=3)) %>%
    dplyr::select(year, disease,median,percentile025,percentile975, scen, title1, age, sex, purpose, title1) %>% mutate(age=ifelse(age=="all", "all age groups", age),
           sex=ifelse(sex=="all", "females and males", sex), 
           sex=ifelse(sex=="male", "males", sex), 
           sex=ifelse(sex=="female", "females", sex)) %>%
                filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
  ggplot(aes(x=year, y=median, colour = disease)) +
    # geom_ribbon(aes(ymin=percentile025,ymax=percentile975),fill="#24C9AC",alpha=0.5) + # AUO teal, removed UI to graph all diseases in one graph
    geom_smooth(se=FALSE) + #or geom_line
    # facet_grid(rows=vars(disease), scales = "free_y") +
    scale_y_continuous(
      name = waiver(),
      breaks = waiver(),
      minor_breaks = NULL,
      n.breaks = 5,
      labels = waiver()) +
    scale_x_continuous(limits=c(0,85),breaks=seq(0,80,10), expand=c(0,0)) +
    labs(x="Years since scenario commenced", y="Deaths prevented") +
    theme_classic() +
    theme(legend.position = "none") +
    theme(plot.title=element_text(size=16, hjust = 0.5),
          plot.subtitle = element_text(size=14, hjust = 0.5),
          axis.text=element_text(size=12),
          axis.title=element_text(size=12),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=12),
          legend.key = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(size=12)) + 
          geom_hline(yintercept = 0, color = "black")})

```

# Health-adjusted life years

### Health-adjusted life years gained per year of simulation for the modelled population.

```{r}

renderPlot({as_tibble(output_df_agg_all) %>%
    filter(measure=="Lwx" & scenario== "diff") %>%
    dplyr::select(year,median,percentile025,percentile975, title1, age, sex, purpose) %>%
    mutate(age=ifelse(age=="all", "all age groups", age),
           sex=ifelse(sex=="all", "females and males", sex), 
           sex=ifelse(sex=="male", "males", sex), 
           sex=ifelse(sex=="female", "females", sex))  %>%
    arrange(year) %>%
    ungroup() %>%
    # year 84 (the last year, has a weird uptick, removing for now)
    filter(year<=83) %>% 
    filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
  ggplot(aes(x=year, y=median)) +
    geom_smooth(se=FALSE) +
    scale_y_continuous(
      name = waiver(),
      breaks = waiver(),
      minor_breaks = NULL,
      n.breaks = 5,
      labels = waiver()) +
    scale_x_continuous(limits=c(0,85),breaks=seq(0,80,10), expand=c(0,0)) +
    labs(x="Years since scenario commenced", y="Health-adjusted life years") +
    # subtitle =  paste0(unique(tmpPlot$population)) +
   theme_classic() +
    labs(x="Years since scenario commenced", y="Change in health-adjusted life years")+
    theme(plot.title=element_text(size=16, hjust = 0.5),
          plot.subtitle = element_text(size=14, hjust = 0.5),
          axis.text=element_text(size=12),
          axis.title=element_text(size=12),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=12),
          legend.key = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(size=12))})
```

# Life years

### Life years gained per year of simulation for the modelled population.

```{r}

renderPlot({as_tibble(output_df_agg_all) %>%
    filter(measure=="Lx" & scenario== "diff") %>%
    dplyr::select(year,median,percentile025,percentile975, title1, age, sex, purpose) %>%
    mutate(age=ifelse(age=="all", "all age groups", age),
           sex=ifelse(sex=="all", "females and males", sex), 
           sex=ifelse(sex=="male", "males", sex), 
           sex=ifelse(sex=="female", "females", sex))  %>%
    arrange(year) %>%
    ungroup() %>%
    # year 84 (the last year, has a weird uptick, removing for now)
    filter(year<=83) %>% 
    filter(age %in% input$age) %>%
                filter(sex %in% input$sex) %>%
                filter(title1 %in% input$scenario) %>%
                filter(purpose %in% input$purpose) %>%
  ggplot(aes(x=year, y=median)) +
    geom_smooth(se=FALSE) +
    scale_y_continuous(
      name = waiver(),
      breaks = waiver(),
      minor_breaks = NULL,
      n.breaks = 5,
      labels = waiver()) +
    scale_x_continuous(limits=c(0,85),breaks=seq(0,80,10), expand=c(0,0)) +
    labs(x="Years since scenario commenced", y="Life years") +
    # subtitle =  paste0(unique(tmpPlot$population)) +
   theme_classic() +
    labs(x="Years since scenario commenced", y="Change in health-adjusted life years")+
    theme(plot.title=element_text(size=16, hjust = 0.5),
          plot.subtitle = element_text(size=14, hjust = 0.5),
          axis.text=element_text(size=12),
          axis.title=element_text(size=12),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          legend.justification = c(1,1),
          legend.title = element_blank(),
          legend.text = element_text(size=12),
          legend.key = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(size=12))})
    
```

# About

This application is in support of the article "Shifting car travel to active modes to improve population health and achieve transport goals: a simulation study" published in xx. The article quantified the health impacts of active transport scenarios for Melbourne.
